Embracing the Laws of Physics: \\ Reversible Models of Computation

### 2017-12-06 survey/survey.tex
next few paragraphs.

\jc{wavefront}
All programs/proofs/deductions are  
isomorphisms/equivalences/deformations.

Take it as a fundamental principle; avoid detour via irreversible
functions and current models of computations.

Starting from a physical perspective, programs naturally emerge as
reversible transformations on data. In some sense, there is nothing
more to say about programs per se. The focus is on the data. Depending
on the space we use to represent the data, the corresponding notion of
deformation on these spaces will define the appropriate notion of
programs. In the following three sections, we consider three natural
classes of data and explore the corresponding notion of reversible
programs.

In the quantum case, this is clear. Say something about topological
quantum computing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data I: Finite Sets}

>>>>

We will take ``the same'' as a fundamental principle. We want
to then derive what it means for data, programs, program
transformations, as well as proofs / deductions, to be ``the
same'' -- in an manner consistent with preservation of 
information.

This stands in stark contrast with most current approacheds to
reversible computation, which start from current models of
computation involving irreversible operations and try to find
various ways to \emph{patch things up} so as to be reversible.

The reader will have surely noticed how we have studiously avoided
defining ``the same''. Because there are many such notions, we
also need to walk our way through them to find the one which is
``just right''. We can draw an analogy with topology: in topology,
all point sets can always be equipped with either the discrete or
the indiscrete topology, but both of these extremes are rarely
useful. We will develop our working notion of ``sameness''
as we go through the various components that make up a
programming language.

Starting from a physical perspective, whatever our notion of data
is, we will be interested in programs as representating transformations
of that data which are reversible. In other words, we want our
programs-as-transformations to ``play well'' with the inherent notion
of ``sameness'' that our data will carry. Thus we need to start by
looking at what structure our data has, and that will help us define
an appropriate notion of program. Of course, when programs themselves
are data, things do get more complicated.  In the following sections,
we will look at different natural classes of data, and explore the
correspond notion of reversible programs.

\amr{In the quantum case, this is clear. Say something about topological
quantum computing.}

\jc{wavefront}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data I: Finite Sets}
